Manuscript

### 2015-03-09 inst/simulation/manuscript.Rmd
...

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
allScenarios <- c(scenarios, paste0(scenarios, "combined"), paste0(scenarios, "10days"))
modelNames <- c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "SmoothModelMean-nbinomial-ar1", "FMPModel")
iterations <- as.integer(c(1,1,1,1,3,1))
populationSizeOverEstimate <- 1e999

prettyScenario <- function(scenario) substr(scenario, 1, 1)
prettyModelName <- function(x) {
  lookup <- data.frame(x=modelNames, y=c("ST", "VT", "T", "FMP"), stringsAsFactors=F)
  lookup$y[match(x, lookup$x)]
}

@@

getPopulationSizes <- function(scenarios, modelNames, populationSizeOverEstimate=Inf) {
  study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
  validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
  populationSizes <- validation$getValidationTemporalPopulationSizes(scenarios=scenarios, modelNames=modelNames)
  #validation$summarizePopulationSize(populationSizes)  
}

@@

```{r figure-3-init-2}
registerDoMC(cores=10) # The nodes seem to run out of memory and fail with a bogus error message, so it's better to reduce the number of used cores.
populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4)])
populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4)])
populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4)])
registerDoMC(cores=detectCores())

@@

xlim <- log10(c(50,1200)); ylim <- xlim
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3a1.png")
saveFigure(p, "manuscript-figure-3a1.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3a2.png")
saveFigure(p, "manuscript-figure-3a2.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3a3.png")
saveFigure(p, "manuscript-figure-3a3.pdf", dpi=300)

xlim <- log10(c(500,4000)); ylim <- xlim
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], "b", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b1.png")
saveFigure(p, "manuscript-figure-3b1.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], "b", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b2.png")
saveFigure(p, "manuscript-figure-3b2.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], "b", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b3.png")
saveFigure(p, "manuscript-figure-3b3.pdf", dpi=300)

xlim <- log10(c(5,1200)); ylim <- xlim
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], "c", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3c1.png")
saveFigure(p, "manuscript-figure-3c1.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], "c", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3c2.png")
saveFigure(p, "manuscript-figure-3c2.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], "c", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3c3.png")
saveFigure(p, "manuscript-figure-3c3.pdf", dpi=300)
```

```{r figure-3sample}
xlim <- log10(c(10,800)); ylim <- xlim
logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
logPopulationSizesSample$modelName <- prettyModelName(logPopulationSizesSample$modelName)
logPopulationSizesSample <- rbind(
  subset(logPopulationSizesSample, modelName=="ST"),
  subset(logPopulationSizesSample, modelName=="T"),
  subset(logPopulationSizesSample, modelName=="FMP"))

@@

p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T","FMP")), modelName=c("ST","T","FMP"), "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3sample.png", height=7.87*1.2/2) # draft quality
saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87*1.2/2, dpi=300)
```

```{r table-1abc}
getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, full=F, file, ...) {
  library(plyr)
  x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
    failed <- (1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x) * 100)
      #(sum(x$Estimated < overEstimate[1] | x$Estimated > overEstimate[2]) / nrow(x)) * 100)
    x <- cutoffPopulationSizes(x, overEstimate)
    #subset(x, Estimated <= overEstimate)
    #if (x$modelName[1] == "FMPModel") failed <- NA
    #if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, SlopeSD=NA, R2=NA, Failed=failed, EstTrue=NA, EstTrueSD=NA, EstTrueMin=NA, EstTrueMax=NA, N=NA))
    y <- lm(Estimated~Observed, x)
    prop.mean <- mean(log(x$Estimated/x$Observed))
    prop.sd <- sd(log(x$Estimated/x$Observed))
    prop.min <- quantile(log(x$Estimated/x$Observed), .05)
    prop.max <- quantile(log(x$Estimated/x$Observed), .95)
    slope.sd <- coef(summary(y))[2,2]
    return(data.frame(Intercept=coef(y)[1], Slope=coef(y)[2], SlopeSD=slope.sd, R2=summary(y)$r.squared, EstTrue=prop.mean, EstTrueSD=prop.sd, EstTrueMin=prop.min, EstTrueMax=prop.max, Failed=failed, N=nrow(x)))
  }, overEstimate=overEstimate)

@@

```{r figure-4-init}
getPopulationCorrelations <- function(scenarios, modelNames) {
  study <- getStudy(scenario="A")
  validation <- Validation(study=study, populationSizeOverEstimate=Inf)
  populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=scenarios, modelNames=modelNames)
  return(populationCorrelations)
}

@@

```{r figure-4abc}
pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1])
pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])

@@

```{r figure-5-init}
getCIProportions <- function(scenarios, modelNames, panel, scenarioNames, cutoff=Inf) {
  study <- getStudy(scenario="A")
  validation <- Validation(study=study, populationSizeOverEstimate=cutoff)
  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames, probs=c(.25, .75), probsName="50%")
  credibilityIntervals <- rbind(a, b)
  credibilityIntervals$panel <- panel
  credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=length(modelNames))
  return(credibilityIntervals)  
}

plotCIProportions <- function(credibilityIntervals) {
  modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
  credibilityIntervals$CI <- credibilityIntervals$Interval
  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=CI, colour=modelName)) +
    geom_hline(aes(yintercept=.5), color="grey90") +
    geom_hline(aes(yintercept=.95), color="grey90") +
    geom_point() +
    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray"), labels=modelLabels) +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
    xlab("Scenario") + ylab("Proportion") +
    facet_wrap(~panel) + ylim(0, 1) +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          plot.margin=unit(c(0, 0, 0, 0), units="line"),
          #legend.background=element_rect(fill="transparent"),
          legend.margin=unit(-2, units="line"),
          panel.margin=unit(0, units="line"),
          legend.position="bottom",
          legend.box="horizontal")
  return(p)
}
```

```{r figure-5abc}
cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[1:3], panel="a", scenarioNames=scenarios, cutoff=cutoff)
cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1:3], panel="b", scenarioNames=scenarios, cutoff=10*cutoff)
cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1:3], panel="c", scenarioNames=scenarios, cutoff=cutoff)

cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
p <- plotCIProportions(cip)
saveFigure(p, "manuscript-figure-5.png", height=7.87/2)
```

>>>>

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
allScenarios <- c(scenarios, paste0(scenarios, "combined"), paste0(scenarios, "10days"))
modelNames <- c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "SmoothModelMean-nbinomial-ar1", "FMPModel", "SmoothModelMean-nbinomial-ar1-priors1")
iterations <- as.integer(c(1,1,1,1,3,1))
#populationSizeOverEstimate <- 1e999

prettyScenario <- function(scenario) substr(scenario, 1, 1)
prettyModelName <- function(x) {
  lookup <- data.frame(x=modelNames, y=c("ST", "VT", "T1", "FMP", "T2"), stringsAsFactors=F)
  lookup$y[match(x, lookup$x)]
}

@@

getPopulationSizes <- function(scenarios, modelNames, populationSizeCutoff=Inf) {
  study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
  validation <- Validation(study=study, populationSizeCutoff=populationSizeCutoff)
  populationSizes <- validation$getValidationTemporalPopulationSizes(scenarios=scenarios, modelNames=modelNames)
  #validation$summarizePopulationSize(populationSizes)  
}

@@

```{r figure-3-init-2}
populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)])
populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4,5)])
populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])

@@

xlim <- log10(c(50,1200)); ylim <- xlim
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3a1.png")
saveFigure(p, "manuscript-figure-3a1.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3a2.png")
saveFigure(p, "manuscript-figure-3a2.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[4]), modelName=modelNames[4], "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3a3.png")
saveFigure(p, "manuscript-figure-3a3.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[5]), modelName=modelNames[5], "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3a4.png")
saveFigure(p, "manuscript-figure-3a4.pdf", dpi=300)

xlim <- log10(c(500,4000)); ylim <- xlim
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], "b", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b1.png")
saveFigure(p, "manuscript-figure-3b1.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], "b", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b2.png")
saveFigure(p, "manuscript-figure-3b2.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[4]), modelName=modelNames[4], "b", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b3.png")
saveFigure(p, "manuscript-figure-3b3.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[5]), modelName=modelNames[5], "b", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b4.png")
saveFigure(p, "manuscript-figure-3b4.pdf", dpi=300)

xlim <- log10(c(5,1200)); ylim <- xlim
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], "c", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3c1.png")
saveFigure(p, "manuscript-figure-3c1.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], "c", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3c2.png")
saveFigure(p, "manuscript-figure-3c2.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[4]), modelName=modelNames[4], "c", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3c3.png")
saveFigure(p, "manuscript-figure-3c3.pdf", dpi=300)
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[5]), modelName=modelNames[5], "c", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3c4.png")
saveFigure(p, "manuscript-figure-3c4.pdf", dpi=300)
```

```{r figure-3sample}
xlim <- log10(c(10,800)); ylim <- xlim
logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
logPopulationSizesSample$modelName <- prettyModelName(logPopulationSizesSample$modelName)
logPopulationSizesSample <- rbind(
  subset(logPopulationSizesSample, modelName=="ST"),
  subset(logPopulationSizesSample, modelName=="T1"),
  subset(logPopulationSizesSample, modelName=="T2"),
  subset(logPopulationSizesSample, modelName=="FMP"))

@@

p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3sample.png", height=7.87*1.2/2) # draft quality
saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87*1.2/2, dpi=300)
```

```{r table-1abc}
getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, full=F, file, ...) {
  library(plyr)
  x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
    failed <- (1 - nrow(cutoffPopulationSizes(x, overEstimate)) / nrow(x)) * 100
      #(sum(x$Estimated < overEstimate[1] | x$Estimated > overEstimate[2]) / nrow(x)) * 100)
    x <- cutoffPopulationSizes(x, overEstimate)
    #subset(x, Estimated <= overEstimate)
    #if (x$modelName[1] == "FMPModel") failed <- NA
    #if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, SlopeSD=NA, R2=NA, Failed=failed, EstTrue=NA, EstTrueSD=NA, EstTrueMin=NA, EstTrueMax=NA, N=NA))
    y <- lm(Estimated~Observed, x)
    prop.mean <- mean(log(x$Estimated/x$Observed))
    prop.sd <- sd(log(x$Estimated/x$Observed))
    prop.min <- quantile(log(x$Estimated/x$Observed), .05)
    prop.max <- quantile(log(x$Estimated/x$Observed), .95)
    slope.sd <- coef(summary(y))[2,2]
    return(data.frame(Intercept=coef(y)[1], Slope=coef(y)[2], SlopeSD=slope.sd, R2=summary(y)$r.squared, EstTrue=prop.mean, EstTrueSD=prop.sd, EstTrueMin=prop.min, EstTrueMax=prop.max, Failed=failed, N=nrow(x)))
  }, overEstimate=overEstimate)

@@

```{r figure-4-init}
getPopulationCorrelations <- function(scenarios, modelNames, cutoff=Inf) {
  study <- getStudy(scenario="A")
  validation <- Validation(study=study, populationSizeOverEstimate=cutoff)
  populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=scenarios, modelNames=modelNames)
  return(populationCorrelations)
}

@@

```{r figure-4abc}
pc1 <- getPopulationCorrelations(scenarios=scenarios, modelNames=modelNames[1], cutoff=cutoff)
pc2 <- getPopulationCorrelations(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1], cutoff=cutoff)
pc3 <- getPopulationCorrelations(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1], cutoff=cutoff)

@@

```{r figure-5-init}
getCIProportions <- function(scenarios, modelNames, panel, scenarioNames, cutoff=Inf) {
  study <- getStudy(scenario="A")
  validation <- Validation(study=study, populationSizeCutoff=cutoff)
  a <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames)
  b <- validation$getValidatedCredibilityIntervalsProportions(scenarios=scenarios, modelNames=modelNames, probs=c(.25, .75), probsName="50%")
  credibilityIntervals <- rbind(a, b)
  credibilityIntervals$panel <- panel
  credibilityIntervals$scenario <- rep(scenarioNames, times=2, each=length(modelNames))
  return(credibilityIntervals)  
}

plotCIProportions <- function(credibilityIntervals) {
  modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
  credibilityIntervals$CI <- credibilityIntervals$Interval
  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=CI, colour=modelName)) +
    geom_hline(aes(yintercept=.5), color="grey90") +
    geom_hline(aes(yintercept=.95), color="grey90") +
    geom_point() +
    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray", "gray"), labels=modelLabels) +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
    xlab("Scenario") + ylab("Proportion") +
    facet_wrap(~panel) + ylim(0, 1) +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          plot.margin=unit(c(0, 0, 0, 0), units="line"),
          #legend.background=element_rect(fill="transparent"),
          legend.margin=unit(-2, units="line"),
          panel.margin=unit(0, units="line"),
          legend.position="bottom",
          legend.box="horizontal")
  return(p)
}
```

```{r figure-5abc}
cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="a", scenarioNames=scenarios, cutoff=cutoff)
cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="b", scenarioNames=scenarios, cutoff=10*cutoff)
cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="c", scenarioNames=scenarios, cutoff=cutoff)

cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
p <- plotCIProportions(cip)
saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)
saveFigure(p, "manuscript-figure-5.pdf", height=7.87/2, dpi=300)
```
